Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae

### 2021-03-23 thesis_chapter_3.Rmd
grammer error in 3.5 title
changed mean traj plots

---
title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
author: 
- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape]
output:
  word_document:
    reference_docx: markdown/frontiers_template.docx
  html_document:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
bibliography: references.bib
csl: markdown/frontiers.csl.txt
abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalage are one of the most important natural dispersal mechanisms, and are regarded as the tumble-weed of the ocean. Despite the many studies that exist, the relative role of wind and surface currents in macrolage trajectory is still unknown. Past studies also focus on kelp-rafts, which can vary greatly in size making it difficult to form a concensus as to the relative role of wind versus surface currents. In addition, past studies have not considered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam. The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagae trajectory and accumulation. This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain. The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumulation zones for solitary floating macroalage."
---

@@

Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown although the role of wind has been recognised as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterrey Bay. The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents, however the relative importance of wind and was not clear. In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter. Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating bull kelp *Durvillaea antarctica*. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For example, a study by @fraser2011 on the rafting capabilities of *D. antarctica* used a combination of population genetics and relative age estimate of goosebarnacles attached to the raft. The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate. Genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011]. Past research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008]. The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag [@breivik2011; @griffin2017]. In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object. If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017]. Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.

@@

p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label

@@

Min_Kfloat_H100_W00$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H100_W00$velocity <- with(Min_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Min_Kfloat_H095_W05$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H095_W05$velocity <- with(Min_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Min_Kfloat_H090_W10$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H090_W10$velocity <- with(Min_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Min_Kfloat_H085_W15$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H085_W15$velocity <- with(Min_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Mean_Kfloat_H100_W00$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_Kfloat_H100_W00$velocity <- with(Mean_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Mean_KFloat_H095_W05$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H095_W05$velocity <- with(Mean_KFloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Mean_KFloat_H090_W10$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H090_W10$velocity <- with(Mean_KFloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Mean_KFloat_H085_W15$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H085_W15$velocity <- with(Mean_KFloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Max_Kfloat_H100_W00$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H100_W00$velocity <- with(Max_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Max_Kfloat_H095_W05$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H095_W05$velocity <- with(Max_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Max_Kfloat_H090_W10$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H090_W10$velocity <- with(Max_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))

@@

Max_Kfloat_H085_W15$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H085_W15$velocity <- with(Max_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))

@@

waterdrag_plot <- ggplot()+
  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'blue')+
  geom_point(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), shape = 20)+
  geom_path(data = Mean_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Max_Kfloat_H100_W00_mean, aes(x = lon, y = lat), color = 'red')+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

wwdrag_plot <- ggplot()+
  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
  geom_path(data = p_RK4_mean, aes(x = lon, y = lat), color = 'black')+
  geom_path(data = Min_Kfloat_H095_W05_mean, aes(x = lon, y = lat), color = 'grey')+
  geom_path(data = Min_Kfloat_H090_W10_mean, aes(x = lon, y = lat), color = 'green')+
  geom_path(data = Min_Kfloat_H085_W15_mean, aes(x = lon, y = lat), color = 'pink')+
  geom_path(data = Mean_KFloat_H095_W05_mean, aes(x = lon, y = lat), color = 'yellow')+
  geom_path(data = Mean_KFloat_H090_W10_mean, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Mean_KFloat_H085_W15_mean, aes(x = lon, y = lat), color = 'red')+
  geom_path(data = Max_Kfloat_H095_W05_mean, aes(x = lon, y = lat), color = 'blue')+
  geom_path(data = Max_Kfloat_H090_W10_mean, aes(x = lon, y = lat), color = 'orange')+
  geom_path(data = Max_Kfloat_H085_W15_mean, aes(x = lon, y = lat), color = 'red')+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Comparison of mean trajectories. Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag. In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10. In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13. The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO")
```

@@

heatmap_RK4 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H100_W00 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H100W100") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H095_W05 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  
  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  #ggtitle("H095W05") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H090_W10 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  
  scale_fill_continuous(type = "gradient",limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H090W10") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H085_W15 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H085W15") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_Kfloat_H100_W00 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H095_W05 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient", limits = c(1, 13000)) +
  geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

>>>>

---
title: 'Numerical experiment investigating the influence of drag on trajectory patterns of floating macroalgae'
author: 
- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape]
output:
  word_document:
    reference_docx: markdown/frontiers_template.docx
  html_document:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
bibliography: references.bib
csl: markdown/frontiers.csl.txt
abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalgae are one of the most important natural dispersal mechanisms, and are regarded as the tumble-weed of the ocean. Despite the many studies that exist, the relative role of wind and surface currents in macroalgae trajectory is still unknown. Past studies also focus on kelp-rafts, which can vary greatly in size making it difficult to form a concensus as to the relative role of wind versus surface currents. In addition, past studies have not considered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam. The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagae trajectory and accumulation. This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain. The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumulation zones for solitary floating macroalage."
---

@@

Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown although the role of wind has been recognised as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterrey Bay. The results showed that kelp rafts, with little surface area, exposed to the wind were largely driven by a combination of wind and surface currents, however the relative importance of wind and was not clear. In addition, the trajectories of tagged kelp were more consistent with the formation of eddies during winter. Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example, in the sub-Antarctic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating bull kelp *Durvillaea antarctica*. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For example, a study by @fraser2011 on the rafting capabilities of *D. antarctica* used a combination of population genetics and relative age estimate of goosebarnacles attached to the raft. The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate. Genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory; however, this was only inferred from the genetic results and local climatology data [@fraser2011]. Past research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [@onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008]. The 'sail' area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag [@breivik2011; @griffin2017]. In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object. If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017]. Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.

@@

p_RK4$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label
p_RK4_mean$type <- "1" # add extra column 'type' to dataframe and assigned 'RK4' label

@@

Min_Kfloat_H100_W00$type <- "Min_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H100_W00$velocity <- with(Min_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H100_W00_mean$type <- "2" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Min_Kfloat_H095_W05$type <- "Min_Kfloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H095_W05$velocity <- with(Min_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H095_W05_mean$type <- "3" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Min_Kfloat_H090_W10$type <- "Min_Kfloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H090_W10$velocity <- with(Min_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H090_W10_mean$type <- "4" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Min_Kfloat_H085_W15$type <- "Min_Kfloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
Min_Kfloat_H085_W15$velocity <- with(Min_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
Min_Kfloat_H085_W15_mean$type <- "5" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_Kfloat_H100_W00$type <- "Mean_Kfloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_Kfloat_H100_W00$velocity <- with(Mean_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_Kfloat_H100_W00_mean$type <- "6" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_KFloat_H095_W05$type <- "Mean_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H095_W05$velocity <- with(Mean_KFloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H095_W05_mean$type <- "7" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_KFloat_H090_W10$type <- "Mean_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H090_W10$velocity <- with(Mean_KFloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H090_W10_mean$type <- "8" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Mean_KFloat_H085_W15$type <- "Mean_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
Mean_KFloat_H085_W15$velocity <- with(Mean_KFloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
Mean_KFloat_H085_W15_mean$type <- "9" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H100_W00$type <- "Max_KFloat_H100_W00" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H100_W00$velocity <- with(Max_Kfloat_H100_W00, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H100_W00_mean$type <- "10" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H095_W05$type <- "Max_KFloat_H095_W05" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H095_W05$velocity <- with(Max_Kfloat_H095_W05, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H095_W05_mean$type <- "11" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H090_W10$type <- "Max_KFloat_H090_W10" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H090_W10$velocity <- with(Max_Kfloat_H090_W10, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H090_W10_mean$type <- "12" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

Max_Kfloat_H085_W15$type <- "Max_KFloat_H085_W15" # add extra column 'type' to dataframe and assigned 'kelp' label
Max_Kfloat_H085_W15$velocity <- with(Max_Kfloat_H085_W15, sqrt(kelpvelU^2 + kelpvelV^2))
Max_Kfloat_H085_W15_mean$type <- "13" # add extra column 'type' to dataframe and assigned 'kelp' label

@@

total_mean_water <- rbind(p_RK4_mean, Min_Kfloat_H100_W00_mean, Mean_Kfloat_H100_W00_mean, Max_Kfloat_H100_W00_mean)
total_mean_ww <- rbind(p_RK4_mean, Min_Kfloat_H095_W05_mean, Min_Kfloat_H090_W10_mean, Min_Kfloat_H085_W15_mean, Mean_KFloat_H095_W05_mean, Mean_KFloat_H090_W10_mean, Mean_KFloat_H085_W15_mean, Max_Kfloat_H095_W05_mean, Max_Kfloat_H090_W10_mean, Max_Kfloat_H085_W15_mean)

total_mean_water$type <- factor(total_mean_water$type, levels=c("1", "2", "6", "10"))
total_mean_ww$type <- factor(total_mean_ww$type, levels=c("1", "3", "4", "5", "7", "8", "9", "11", "12", "13"))

waterdrag_plot <- ggplot()+
  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
  geom_path(data = total_mean_water, aes(x = lon, y = lat, colour = type))+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  scale_colour_discrete("Type") +  
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

wwdrag_plot <- ggplot()+
  #geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID), size = 1.0, colour = "black", fill = NA) +
  geom_path(data = total_mean_ww, aes(x = lon, y = lat, colour = type))+
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  scale_colour_discrete("Type") +  
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_bw()

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Comparison of mean trajectories. Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag. The reader is referred to tables 1 and 2 for details pertaining to each scenario."}

ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO", ncol = 1, nrow = 2)

@@

heatmap_RK4 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_RK4, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = p_RK4, bins = 80, aes(x = lon, y = lat), legend = FALSE) +  
  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
  geom_path(data = p_RK4_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  ggtitle("") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H100_W00 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
  scale_fill_continuous(type = "gradient",limits = c(1, 1000)) +
  geom_path(data = Min_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H100W100") +
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H095_W05 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  
  scale_fill_continuous(type = "gradient",limits = c(1, 1000)) +
  geom_path(data = Min_Kfloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  #ggtitle("H095W05") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H090_W10 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H090_W10, bins = 80, aes(x = lon, y = lat)) +  
  scale_fill_continuous(type = "gradient",limits = c(1, 1000)) +
  geom_path(data = Min_Kfloat_H090_W10_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H090W10") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Min_Kfloat_H085_W15 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Min_Kfloat_H085_W15, bins = 80, aes(x = lon, y = lat)) +  
  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
  geom_path(data = Min_Kfloat_H085_W15_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  #ggtitle("H085W15") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_Kfloat_H100_W00 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = p_EE, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_Kfloat_H100_W00, bins = 80, aes(x = lon, y = lat)) +  
  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
  geom_path(data = Mean_Kfloat_H100_W00_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

heatmap_Mean_KFloat_H095_W05 <- ggplot() +
  geom_polygon(data = south_africa_coast_hi_res, aes(x = lon, y = lat, group = PID),
               size = 1.0, colour = "black", fill = NA) +
  #geom_bin2d(data = k_app1, bins = 80, aes(x = lon, y = lat)) +
  geom_hex(data = Mean_KFloat_H095_W05, bins = 80, aes(x = lon, y = lat)) +  scale_fill_continuous(type = "gradient", limits = c(1, 1000)) +
  geom_path(data = Mean_KFloat_H095_W05_mean,aes(x = lon, y = lat, color = 'red'), show.legend = FALSE) +
  coord_cartesian(xlim = c(14, 20), ylim = c(-35, -30)) + 
  xlab("Longtitude") + 
  ylab("Latitude") + 
  guides(fill=guide_legend(title="Number of particles")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
